---
title: "MCnebula workflow for LC-MS/MS dataset analysis"
bibliography: '`r system.file("extdata", "library.bib", package = "utils.tool")`'
csl: '`r system.file("extdata", "styles.csl", package = "utils.tool")`'
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: true
    toc_depth: 3
    latex_engine: xelatex
    df_print: kable
---

# MCnebula workflow

## Introduction

We know that the analysis of untargeted LC-MS/MS dataset generally begin with feature detection.
It detects 'peaks' as features in MS^1^ data.
Each feature may represents a compound, and assigned with MS^2^ spectra.
The MS^2^ spectra was used to find out the compound identity.
The difficulty lies in annotating these features to discover their compound identity, mining out meaningful information, so as to serve further biological research.
Herein, a classified visualization method, called MCnebula, was used for addressing this difficulty.
MCnebula utilizes the state-of-the-art computer prediction technology, SIRIUS workflow (SIRIUS, ZODIAC, CSI:fingerID, CANOPUS)[@2019; @2009; @2015; @2020a; @2015a; @2021a], for compound formula prediction, structure retrieve and classification prediction.
MCnebula integrates an abundance-based class selection algorithm into features annotation: depending on the user, MCnebula parses chemical classes with more or less features in the dataset (the abundance of classes), visualizes them, and displays the feautres they involved;
these classes can be dominant structural classes or sub-structural classes.
With MCnebula, we can switch from untargeted to targeted analysis, focusing precisely on the compound or chemical class of interest to the researcher.  

&ensp;&ensp; 
MCnebula primarily performs an abundance-based class selection algorithm before visualization.
MCnebula tend to filter out those classes with too large scope (e.g., possibly be 'Lipids and lipid-like molecules' but data dependent) or too sparse compounds (data depend).
We termed these summarised classes as nebula-index.
To begin with, like classical feature-based molecular networking (FBMN) pattern [@2020d], features make up the initial network, which we termed parent-nebula.
Subsequently, according to nebula-index and the posterior probability of classes prediction (PPCP) of features [@2021a], nodes or edges from parent-nebula are divided into sub-networks.
We termed these sub-networks as child-nebulae and their names, termed nebula-name, are in line with the classes name within nebulae-index.
The nebula-names contained the sub-structural or dominant-structural characteristics for features within child-nebulae.
Collectively, all the network and sub-networks termed multi-chemical nebulae.
In general, parent-nebula is too informative to show, so child-nebulae was used to dipict the abundant classes of metabolites in a grid panel, intuitively.
In a bird's eye view of child-nebulae, we can obtain many characteristics of features, involving classes distribution, structure identified accuracy, as well as spectral similarity within classes.

## R and other softs Setup

### R setup

```{r setup}
library(MCnebula)
library(dplyr)
library(ggplot2)
library(ggraph)
library(grid)
```

### Others setup

The prerequisite soft for MCnebula is [SIRIUS 4](https://bio.informatik.uni-jena.de/software/sirius/).  

&ensp;&ensp; 
MZmine 2 (version 2.53) (<https://github.com/mzmine/mzmine2>) were utilized for LC-MS/MS data processing.

&ensp;&ensp; 
Before mass spectrometry data been processed, the raw data should be converted to .mzML or .mzXML file in most cases.
ProteoWizard (<https://proteowizard.sourceforge.io/>) were implemented for the conversion.

## Data preprocessing

### Raw data processing

For MZmine2 processing, an XML batch file outlined the example parameters for waters Qtof could be find in <https://github.com/Cao-lab-zcmu/research-supplementary>.

### SIRIUS computation workflow

The computational parameters of SIRIUS CLI tools were set as following:

```{r, eval = F}
mgf.path <- "my.mgf"
system(
  paste0(
    "sirius -i ",
    mgf.path,
    " -o test --maxmz 800 formula -c 50 zodiac structure canopus"
    )
)
```

### MCnebula processing

@start

## Guidance

### Details of workflow

An end-to-end analysis of MCnebula workflow from sample to multi-chemical-nebulae is show in figure \@ref(fig:workflow)

```{r workflow, echo = F, fig.cap = "workflow"}
file <- system.file("extdata", "workflow.png", package = "utils.tool")
png <- png::readPNG(file)
grid::grid.raster(png)
```

## Information

All output files:

```{r, eval = F}
list.files(paste0(.MCn.output, "/", .MCn.results), recursive = T)
```

Other information:

```{r, echo = T}
sessionInfo()
```

## Reference

